\documentclass{article}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{parskip} % so I don't have to use \noindent everywhere
\usepackage{mathtools}
\usepackage[none]{hyphenat} % don't split up words that flow onto next line
\usepackage{microtype,todonotes}
\usepackage[a4paper,text={16.5cm,25.2cm},centering]{geometry}
\usepackage[compact,small]{titlesec}
\setlength{\parskip}{1.2ex}
\setlength{\parindent}{0em}

% For addition and multiplication tables
\usepackage{dcolumn}
\newcolumntype{2}{D{.}{}{2.0}}

% To prevent orphans and widows (https://en.wikipedia.org/wiki/Widows_and_orphans)
\clubpenalty = 10000
\widowpenalty = 10000
\usepackage{kpfonts}
\usepackage[T1]{fontenc}

\title{Abstract Algebra Homework 9}
\author{\textit{Joe Loser}}
\date{April 11, 2016}
\begin{document}
\maketitle

This problem set includes problems $5b, 8, 18c$, and $27$ from section $16.6$.

5b) For the given ring $R$ with an ideal $I$, give an addition and multiplication table for $R/I$.

\underline{Solution}: Recall that 
$$R/I := \{ r+I \, \vert \, r \in R \}.$$

We can easily see that the three elements of $R/I$ are the following:
\begin{align*}
	0 + I &= \{ 0, 3, 6, 9 \} \\
	1 + I &= \{ 1, 4, 7, 10 \} \\
	2 + I &= \{ 2, 5, 8, 11 \}.
\end{align*}

Below is the addition table for $R/I$. Note that it is implicit, but worth noting, that we are talking about the addition of the three cosets here in $R/I$.

\begin{center}
	\begin{tabular}{r|*{4}{2|}}
		+ & 0 & 1 & 2 \\
		\hline
		0 & 0 & 1 & 2 \\ 
		\hline
		1 & 1 & 2 & 0  \\ 
		\hline
		2 & 2 & 0 & 1
	\end{tabular}
\end{center}

Similarly, here is the multiplication table for $R/I$. Again, we are talking about multiplication of the cosets in $R/I$.

\begin{center}
	\begin{tabular}{r|*{4}{2|}}
		* & 0 & 1 & 2 \\
		\hline
		0 & 0 & 0 & 0 \\ 
		\hline
		1 & 0 & 1 & 2  \\ 
		\hline
		2 & 0 & 2 & 1
	\end{tabular}
\end{center}
\qed \\

8) Prove or disprove: The ring $\mathbb{Q}(\sqrt{2}) = \{ a+b\sqrt{2} \, \vert \, a, b \in \mathbb{Q} \}$ is isomorphic to the ring $\mathbb{Q}(\sqrt{3}) = \{ a+b\sqrt{3} \, \vert \, a, b \in \mathbb{Q} \}$.

\underline{Proof}: We will show that $\mathbb{Q}(\sqrt{2}) \not \cong \mathbb{Q}(\sqrt{3})$. To do this, we need to show that no homomorphism from $\mathbb{Q}(\sqrt{2})$ to $\mathbb{Q}(\sqrt{3})$ can be an isomorphism.

Suppose that $\phi : \mathbb{Q}(\sqrt{2}) \to \mathbb{Q}(\sqrt{3})$ is a homomorphism. We will begin by showing that $\phi$ fixes $\mathbb{Z}$ and $\mathbb{Q}$.

We first show that $\phi(1) = 1$ since we do not get this for free by our definition of a ring homomorphism. Let $x \in \mathbb{Z}$. Suppose $\phi(1) = a + b\sqrt{3}$ where $a, b \in \mathbb{Q}$. Then we have that
\begin{align*}
	\phi(x) &= \phi(x \cdot 1) \\
	&= \phi(x) \phi(1) \\
	&= \phi(x \dot 1) \phi(1) \\
	&= \phi(x) \phi(1)^2 \\
	&= \vdots \\
	&= \phi(x) \phi(1)^n.
\end{align*}
So $\phi(x) = \phi(x) \phi(1)^n$. Thus $1 = \phi(1)^n$. Hence $\phi(1) = 1$. 

We now use this result of $\phi(1) = 1$ to show that $\phi$ fixes $\mathbb{Z}$. That is, we extend it to show $\phi(n) = n$ for all $n \in \mathbb{Z}$. To show this, let $n \in \mathbb{Z^+}$. Then
\begin{align*}
	\phi(n) &= \phi \underbrace{(1 + \cdots + 1)}_{n \text {-times}} \\
	&= \underbrace{\phi(1) + \cdots + \phi(1)}_{n \text {-times}} \\
	&= \underbrace{1 + \cdots + 1}_{n \text {-times}} \\
	&= n.
\end{align*}
The proof is similar for $n \in \mathbb{Z^-}$ and we conclude that $\phi(n) = n$ for all $n \in \mathbb{Z}$.

Next, we show that $\phi$ fixes $\mathbb{Q}$. That is, $\phi(y) = y$ for all $y \in \mathbb{Q}$. Let $y = \frac ab$ where $a, b \in \mathbb{Z}$. Then 
\begin{align*}
	\phi(y) &= \phi\big(\frac ab\big) \\
	&= \phi(ab^{-1}) \\
	&= \phi(a) \phi(b^{-1}) \\
	&= \phi(a) \phi(b)^{-1} \\
	&= \frac{\phi(a)}{\phi(b)} \\
	&= \frac ab
\end{align*}
since we just showed $\phi(n) = n$ for all $n \in \mathbb{Z}$ and both $a, b \in \mathbb{Z}$. Thus $\phi(y) = y$ for all $y \in \mathbb{Q}$.

Therefore, if $a + b\sqrt{2} \in \mathbb{Q}(\sqrt{2})$ we see that 
\begin{align*}
	\phi(a + b\sqrt{2}) &= \phi(a) + \phi(b\sqrt{2}) \\
	&= \phi(a) + \phi(b) \phi(\sqrt{2}) \\
	&= a + b \phi(\sqrt{2}).
\end{align*}
So we need to figure out what exactly $\phi(\sqrt{2})$ is.

If $\phi(\sqrt{2}) = c + d\sqrt{3}$ for some $c, d \in \mathbb{Q}$ then what are the possible values of $c$ and $d$? First notice that $\phi(2) = 2$ since $2 \in \mathbb{Z}$. We also know that $2 = (\sqrt{2})^2$. So then 
\begin{align*} 
	\phi(2)  &= \phi \big( (\sqrt{2})^2 \big) \\
	&= \phi(\sqrt{2}) \phi(\sqrt{2}) \\
	&= (c+d\sqrt{3})^2 \\
	&= c^2 + 3d^2 + 2cd\sqrt{3}.
\end{align*}
Thus $2 = c^2 + 3d^2$ and $0 = 2cd\sqrt{3}$. Hence $cd = 0$. So either $c = 0$ or $d = 0$. 

If $c = 0$ then we have $2 = 3d^2$. So $d = \sqrt{\frac 23} \not \in \mathbb{Q}$ which is a contradiction since $d \in \mathbb{Q}$. Similarly, if $d = 0$ then we have that $2 = c^2$. So $c = \sqrt 2 \not \in \mathbb{Q}$ which is a contradiction since $c \in \mathbb{Q}$.

Thus $\phi(\sqrt{2}) = c + d\sqrt{3}$ for $c, d \in \mathbb{Q}$. But we have shown that we cannot find a suitable $c$ or $d \in \mathbb{Q}$ such that this is satisfied. Hence we have shown that there is no isomorphism from $\mathbb{Q}(\sqrt{2})$ to $\mathbb{Q}(\sqrt{3})$. Therefore, as rings, $\mathbb{Q}(\sqrt{2}) \not \cong \mathbb{Q}(\sqrt{3})$. \qed \\

\pagebreak 18c) Let $\phi : R \to S$ be a ring homomorphism. Let $1_R$ and $1_S$ be the identities for $R$ and $S$, respectively. If $\phi$ is onto, show that $\phi(1_R) = 1_S$. 

\underline{Proof}: Let $R$ and $S$ be rings with identities $1_R$ and $1_S$. Suppose $\phi$ is a ring homomorphism from $R$ to $S$.

Let $e = \phi(1_R)$. So $e \in S$. Also let $s \in S$. Since $\phi$ is onto, there exists an element $r \in R$ so that $\phi(r) = s$. Then we have that
\begin{align*}
	es &= \phi(1_R) \phi(r) \\
	&= \phi(1_R \cdot r) \\
	&= \phi(r) \\
	&= s.
\end{align*}
So $es = s$. We now show that that $se = s$ as well. 

\begin{align*}
	se &= \phi(r) \phi(1_R) \\
	&= \phi(r \cdot 1_R) \\
	&= \phi(r) \\
	&= s.
\end{align*}

Therefore $e \in S$ and $es = s = se$ for all $s \in S$. Hence $e$ is a multiplicative identity for the ring $S$. However, $1_S$ is an identity as well. We have shown in class that if a ring has a multiplicative identity, it must be unique. We did this by showing that the set of units in a ring is a group and we know that the identity in a group is unique, hence the $1$ in a ring is unique. Thus, it must be the case that $e = 1_S$. Since we defined $e$ as $e = \phi(1_R)$ we conclude that $\phi(1_R) = 1_S$. \qed \\

27) Let $R$ be a commutative ring. An element $a \in R$ is nilpotent if $a^n = 0$ for some positive integer $n$. Show that the set of all nilpotent elements forms an ideal in $R$. 

\underline{Proof}: Let $N$ be the set of all nilpotent elements, i.e. the nilradical. To show $N$ is an ideal, we need to show

(i) $(N, +) < (R, +)$ \\
(ii) $rx \in N$ and $xr \in N$ for all $r \in R$ and for all $x \in N$.

Clearly $N$ is nonempty as $0 = 0^1 \in N$. Let $x, y \in N$. Then $x^n = y^m = 0$ for some $n, m \in \mathbb{N}$. By the Binomial Theorem we have that
$$ (x+y)^{n+m} = \sum_{k=0}^{n+m} { {n+m}\choose{k} } x^k y^{n+m-k}.$$

Note that if $k \geq n, x^k = 0$. So $x^ky^{n+m-k} = 0$. Similarly, if $k < n, y^{n+m-k} = 0$ so $x^ky^{n+m-k} = 0$. Thus $(x+y)^{n+m} =0$ and hence $(x+y) \in N$. Notice that we needed the fact that $R$ is commutative for this to hold.

To show that additive inverses exist in $N$, note that for $x \in N$ we have 
\begin{align*}
	(-x)^n &= (-1)^n x^n \\
	&= (-1)^n \cdot 0 \\
	&= 0.
\end{align*}
So $-x \in N$. Thus $N$ is an additive subgroup of $R$.

Lastly, let $r \in R$ and $x \in N$. Since $R$ is commutative, we should only need to show $(xr)^n = 0$ and then automatically would we have $(rx)^n = 0$. We show both calculations here nonetheless.
\begin{align*}
	(xr)^n &= r^n x^n \\
	&= r^n \cdot 0 \\
	&= 0.
\end{align*}
Similarly, 
\begin{align*}
	(rx)^n &= x^n r^n \\
	&= 0 \cdot r^n \\
	&= 0.
\end{align*}
So both $rx \in N$ and $xr \in N$.

Since $N$ is an additive subgroup of $R$ and $N$ absorbs $R$ on the left and right, $N$ is an ideal of $R$. \qed \\

\end{document}